---
title: "Time Series Analysis - Day 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

### Packages and libraries installation.

```{r}
#install.packages('forecast', dependencies = TRUE)
library(forecast)
library(ggplot2)
```

Forecast gives usefull tools for time series analysis.

### Time series "Varicelle" analysis.

```{r}
data=read.csv(file="http://eric.univ-lyon2.fr/~jjacques/Download/DataSet/varicelle.csv")
plot(data$x)
varicelle<-ts(data$x,start=c(1931,1),end=c(1972,6),freq=12)
plot(varicelle)
```

We observe a cyclic pattern.

```{r}
mean(varicelle)
```

```{r}
tmp=acf(varicelle, type="cor", plot=TRUE)
ptmp=pacf(varicelle, type="cor", plot=FALSE)
plot (ptmp)
```

The correlogram shows that the data are cyclic. We can see a correlation close to 1 every lag=1 so data looks correlated from one year to the next one.

```{r}
ggseasonplot(varicelle,year.labels= TRUE,year.labels.left=TRUE)
```

The seasonal plot confirm this trend, we can see that every year we have an increase on spring and a decrease on summer.

Let's print the graph year by year instead of month by month.

```{r}
plot(aggregate(varicelle,nfrequency=1))
```

We dont see now a cyclic pattern. However we can see trends corresponding to historical events, like the dramatic decrease during the 2nd world war (due to the decrease of the natality).
